---
output:
  pdf_document: default
---


# Introduction
`weather_features.csv` contains hourly weather data from 2015 to 2018 for 5 major cities in Spain.

```{r}
weather_features <- read.csv("weather_features.csv")
df <- data.frame(
  Column = colnames(weather_features),
  Description = c("datetime index localized to CET", 
                  "name of city (Barcelona, Bilbao, Madrid, Seville, Valencia)",
                  "temperature (K)", 
                  "minimum temperature (K)",
                  "maximum temperature (K)",
                  "pressure (hPa)",
                  "humidity (%)",
                  "wind speed (m/s)",
                  "wind direction (degrees)",
                  "rain in last hour (mm)",
                  "rain in last 3 hours (mm)",
                  "snow in last 3 hours (mm)",
                  "cloud cover (%)",
                  "weather description - code",
                  "weather description - short",
                  "weather description - long",
                  "weather icon")
)
knitr::kable(df, format = "markdown")
```


# Missing Values
```{r}
sapply(weather_features, function(col)
  { sum(sapply(col, function(x) {
    (is.na(x)) + (x == "")
  })) })
```

Data has no blank or NA cells in any of the columns. 
The last 4 columns are non-numerical, and will be dropped. The data for the city of Barcelona has city_name " Barcelona", the extra space will be removed. The datetime index will be switched to date format.
```{r}
weather_features <- weather_features[1:13]
weather_features[weather_features$city_name == " Barcelona",]$city_name = "Barcelona"
weather_features$dt_iso <- as.POSIXct(weather_features$dt_iso)
```

Next, we check that there are no missing rows.  
Since we have hourly data for 3 regular and 1 leap year, we expect there to be $24 \times 365 \times 4 + 24 = 35064$ unique datetime indices.
```{r}
dates_occurences <- table(weather_features$dt_iso)
length(dates_occurences)
```
This matches our expectations.

Since we have 5 cities, we expect each datetime index to appear 5 times. 
```{r}
sum(sapply(dates_occurences, function(date) { date < 5 }))
```
This indicates that we have no missing rows. 


# Partition Based on City
This means a total of $35064 \times 5 = 175320$ rows. However, our data has 178396 rows. This indicates the presence of duplicated rows.

To fix this, we first create separate data frames for each city, then remove duplicates.
```{r}
library(dplyr)
Barcelona <- weather_features[weather_features$city_name == "Barcelona", ] %>% distinct()
Bilbao <- weather_features[weather_features$city_name == "Bilbao", ] %>% distinct()
Madrid <- weather_features[weather_features$city_name == "Madrid", ] %>% distinct()
Seville <- weather_features[weather_features$city_name == "Seville", ] %>% distinct()
Valencia <- weather_features[weather_features$city_name == "Valencia", ] %>% distinct()

dim(Barcelona)
dim(Bilbao)
dim(Madrid)
dim(Seville)
dim(Valencia)
```
